---
title: "ARMA basics"
output: pdf_document
---

```{r setup, include=FALSE}
library(MTS)
library(stats)
library(dplyr)
library(readxl)
library(DataAnalytics)
library(data.table)

```

# Problem 1: ARMA basics

1. ARMA(1,1) model is given by:

$$r_t - \phi_1r_{t-1} = \phi_0 + \epsilon_t - \theta_1\epsilon_{t-1}$$

First order autocorrelation is given by:

$$ \rho_1 = \phi_1 - \theta_1 \frac{\sigma^2_\epsilon}{\gamma_0}$$
Here, $\phi_1 = 0.95$, $\theta_1 = 0.9$, $\sigma^2_{\epsilon} = 0.05^2$. Computation of $\gamma_0$:

$$ \gamma_0 = \sigma^2_{\epsilon} \frac{1+\theta^2_1 - 2\phi_1\theta_1}{1-\phi^2_1}$$
Substituting, we get $\gamma_0 = 0.002564$

Therefore, calculating $\rho_1$, we get $\rho_1 = 0.0725$

2. Second order autocorrelation of $y_t$ is given by: $\rho_2 = \phi_1\rho_1$

Therefore, $\rho_2 = 0.68875$

Ratio of second order to first order autocorrelation = $\phi_1 = 0.95$

This result is expected because ARMA models gets its correlation just like AR(p) models and hence the ratio is just $\phi_1$. This also tells us that the autocorrelations decrease over time if $\phi_1 < 1$.

3. Irrespective of $y_t$, ARMA models have the same mean for the series since it is stationary, and hence

$$ E[y_{t+1}] = E[y_{t+2}] = \frac{\phi_0}{1-\phi_1} = 0$$
4. Given ${\hat{x_t}} = E[y_{t+1}]$. Therefore, ${\hat{x_t}}$ is always a constant that 

```{r}



```