---
title: "Comparison of Returns using CAPM and Real world"
output: pdf_document
---

## Initial Setup

```{r setup, echo=TRUE, inclue=FALSE, results='hide', message=FALSE}
library(dplyr)
library(ggplot2)
library(lubridate)
library(quantmod)
library(xts)
library(psych)
library(IntroCompFinR)
library("readxl")
library(kableExtra)

MSFT <- read_excel('~/Documents/Documents/Investment Codes/lecture6p.xlsx',
                   sheet = "MSFT") 
MSFT$Date <- as.Date(as.character(MSFT$Date))
MSFT = subset(MSFT, select = -c(Volume, Open, High, Low, Close))
MSFT$AdjClose <- as.numeric(MSFT$AdjClose)
MSFT <- xts(MSFT$AdjClose, order.by = MSFT$Date)
INTC <- read_excel('~/Documents/Documents/Investment Codes/lecture6p.xlsx',
                   sheet = "INTC") 
INTC$Date <- as.Date(as.character(INTC$Date))
INTC = subset(INTC, select = -c(Volume, Open, High, Low, Close))
INTC$AdjClose <- as.numeric(INTC$AdjClose)
INTC <- xts(INTC$AdjClose, order.by = INTC$Date)
LUV <- read_excel('~/Documents/Documents/Investment Codes/lecture6p.xlsx',
                   sheet = "LUV") 
LUV$Date <- as.Date(as.character(LUV$Date))
LUV = subset(LUV, select = -c(Volume, Open, High, Low, Close))
LUV$AdjClose <- as.numeric(LUV$AdjClose)
LUV <- xts(LUV$AdjClose, order.by = LUV$Date)
MCD <- read_excel('~/Documents/Documents/Investment Codes/lecture6p.xlsx',
                   sheet = "MCD") 
MCD$Date <- as.Date(as.character(MCD$Date))
MCD = subset(MCD, select = -c(Volume, Open, High, Low, Close))
MCD$AdjClose <- as.numeric(MCD$AdjClose)
MCD <- xts(MCD$AdjClose, order.by = MCD$Date)
JNJ <- read_excel('~/Documents/Documents/Investment Codes/lecture6p.xlsx',
                   sheet = "JNJ") 
JNJ$Date <- as.Date(as.character(JNJ$Date))
JNJ = subset(JNJ, select = -c(Volume, Open, High, Low, Close))
JNJ$AdjClose <- as.numeric(JNJ$AdjClose)
JNJ <- xts(JNJ$AdjClose, order.by = JNJ$Date)
RF <- read_excel('~/Documents/Documents/Investment Codes/lecture6p.xlsx',
                   sheet = "F-F_Research_Data_Factors_daily") 
RF$Date <- as.Date(as.character(RF$Date))
RF = subset(RF, select = -c(SMB, HML, rawdate))
RF$RF <- as.numeric(RF$RF)
RF_p <- 0
for(i in 1:length(RF$RF)) {
  if(i==1) {RF_p[i] <- (1+RF$RF[i]/100)}
  else{RF_p[i] <- RF_p[i-1]*(1+RF$RF[i]/100)}
}
RF <- xts(RF_p, order.by = RF$Date)

MktRF <- read_excel('~/Documents/Documents/Investment Codes/lecture6p.xlsx',
                    sheet ="F-F_Research_Data_Factors_daily") 
MktRF$Date <- as.Date(as.character(MktRF$Date))
MktRF = subset(MktRF, select = -c(SMB, HML, rawdate))
MktRF$MktRF <- as.numeric(MktRF$MktRF)
MktRF_p <- 0
for(i in 1:length(MktRF$MktRF)) {  
  if(i==1) {MktRF_p[i] <- (1+MktRF$MktRF[i]/100)}
  else{MktRF_p[i] <- MktRF_p[i-1]*(1+MktRF$MktRF[i]/100)}
  }
MktRF <- xts(MktRF_p, order.by = MktRF$Date)
nss <- function(t, b0, b1, b2, b3, t1, t2) {
    r <- b0 + b1*((1-exp(-t/t1))/(t/t1)) + (b2*(((1-exp(-t/t1))/(t/t1))-
          (exp(-t/t1)))) +  (b3*(((1-exp(-t/t2))/(t/t2))-(exp(-t/t2))))
    return(r)
}

calculate_bond_price <- function(maturity, par, rate) {
  bond_price <- par/((1+(rate/100))^maturity)
  return(bond_price)
}

```



(1) Construct weekly simple total returns from the price data (use Adj Close to include div- idends). Compute and report the weekly and annualized mean and standard deviation for each stock. Compute the correlation matrix.
```{r 1, echo=TRUE}
MSFT_return <- weeklyReturn(MSFT) 
colnames(MSFT_return) <- "MSFT"
MSFT_sd <- sd(MSFT_return)*sqrt(52)
MSFT_mean <- mean(MSFT_return) * 52
INTC_return <- weeklyReturn(INTC) 
colnames(INTC_return) <- "INTC"
INTC_sd <- sd(INTC_return)*sqrt(52)
INTC_mean <- mean(INTC_return) * 52
LUV_return <- weeklyReturn(LUV) 
colnames(LUV_return) <- "LUV"
LUV_sd <- sd(LUV_return)*sqrt(52)
LUV_mean <- mean(LUV_return) * 52
MCD_return <- weeklyReturn(MCD) 
colnames(MCD_return) <- "MCD"
MCD_sd <- sd(MCD_return)*sqrt(52)
MCD_mean <- mean(MCD_return) * 52
JNJ_return <- weeklyReturn(JNJ) 
colnames(JNJ_return) <- "JNJ"
JNJ_sd <- sd(JNJ_return)*sqrt(52)
JNJ_mean <- mean(JNJ_return) * 52
RF_return <- weeklyReturn(RF) 
colnames(RF_return) <- "RF"
RF_sd <- sd(RF_return)*sqrt(52)
RF_mean <- mean(RF_return) * 52
MktRF_return <- weeklyReturn(MktRF) 
colnames(MktRF_return) <- "MktRF"
MktRF_sd <- sd(MktRF_return)*sqrt(52)
MktRF_mean <- mean(MktRF_return) * 52
Corr_matrix <- cor(data.frame(MSFT_return, INTC_return, LUV_return, 
                               MCD_return, JNJ_return))
Cov_All <- cov(data.frame(MSFT_return, INTC_return, LUV_return, 
                               MCD_return, JNJ_return)) * 52
Cov_MSFT_INTC <- cov(data.frame(MSFT_return, INTC_return)) * 52

data.frame("Stock" = c("MSFT", "INTC", "LUV", "MCD", "JNJ"), 
           "Mean" = c(MSFT_mean, INTC_mean, LUV_mean, MCD_mean, JNJ_mean), 
           "Standard Deviation" =c(MSFT_sd, INTC_sd, LUV_sd, MCD_sd, JNJ_sd))%>%
  kbl() %>%kable_styling()


```


(2) Construct the mean-variance frontier for the Intel-Microsoft combination. Indicate the minimum-variance portfolio and the efficient frontier (the efficient frontier is a set of expected returns - risks that you would want to consider investing in).
```{r s, echo=TRUE}

ef <- efficient.frontier(c(MSFT_mean,INTC_mean), Cov_MSFT_INTC)

# compute global minimum variance portfolio for Microsoft and Intel

gmin.port = globalMin.portfolio(c(MSFT_mean,INTC_mean), Cov_MSFT_INTC)
plot(ef$sd, ef$er, col="blue", pch=16, type = 'l', xlab = "Standard Deviation", 
     ylab = "Expected Return")
points(gmin.port$sd, gmin.port$er, col="red", pch=16, cex=2)
text(gmin.port$sd, gmin.port$er, labels="MIN VARIANCE", pos=4)
lines(ef$sd[1:15], ef$er[1:15], col="green", pch=16, type = 'l')

ef
gmin.port
```


3. Add remaining stocks to the mix. Compute the mean-variance frontier and plot it
on the same chart with the one from the previous question. Indicate the minimum variance portfolio and the efficient frontier. How do they compare to those of the previous question?


```{r 3, echo=TRUE}

ef_all <- efficient.frontier(c(MSFT_mean,INTC_mean,JNJ_mean, LUV_mean, MCD_mean)
                             , Cov_All)

# compute global minimum variance portfolio for all assets
gmin.port.all = globalMin.portfolio(c(MSFT_mean,INTC_mean,JNJ_mean, LUV_mean, 
                                      MCD_mean), Cov_All)
plot(ef_all$sd, ef_all$er, col="blue", pch=16, type = 'l', 
     xlab = "Standard Deviation", ylab = "Expected Return")
points(gmin.port.all$sd, gmin.port.all $er, col="red", pch=16, cex=2)
text(gmin.port.all$sd, gmin.port.all$er, labels="MIN VARIANCE", pos=4)
lines(ef_all$sd[1:15], ef_all$er[1:15], col="green", pch=16, type = 'l')

ef_all
gmin.port.all
```

From the minimum variance portfolio and the efficient frontier, we can observe that:
1. As the number of stocks increased, volatility of the portfolio decreased with almost same amount of the return
2.The efficient frontier has shifted towards the left slightly, indicating lower volatilities




4. Add the riskless asset and construct the tangent portfolio for the Intel-Microsoft case.
Next, construct the tangent portfolio for the full set of stocks. Compare the Sharpe
ratios of the two tangent portfolios.


```{r 4, echo=TRUE}

ef_INTC_MSFT_RF <- efficient.frontier(c(MSFT_mean,INTC_mean), Cov_MSFT_INTC)

# tangency portfolio
tan.port <- tangency.portfolio(c(MSFT_mean,INTC_mean), Cov_MSFT_INTC, RF_mean)

# compute global minimum variance portfolio
gmin.port.INTC_MSFT_RF = globalMin.portfolio(c(MSFT_mean,INTC_mean), 
                                             Cov_MSFT_INTC)
plot(ef_INTC_MSFT_RF$sd, ef_INTC_MSFT_RF$er, col="blue", pch=16, type = 'l', 
     xlab = "Standard Deviation", ylab = "Expected Return", xlim = c(0,0.31), 
     ylim = c(0, 0.27))
points(gmin.port.INTC_MSFT_RF$sd, gmin.port.INTC_MSFT_RF$er, col="red", pch=16,
       cex=2)
lines(ef_INTC_MSFT_RF$sd[1:15], ef_INTC_MSFT_RF$er[1:15], col="green", pch=16, 
      type = 'l')
points(tan.port$sd, tan.port$er, col="magenta", pch=16, cex=2)
text(gmin.port.INTC_MSFT_RF$sd, gmin.port.INTC_MSFT_RF$er, labels="GLOBAL MIN", 
     pos=1)
text(tan.port$sd, tan.port$er, labels="TANGENCY", pos=3)
sr.tan = (tan.port$er - RF_mean)/tan.port$sd
abline(a=RF_mean, b=sr.tan, col="orange", lwd=2)


plot(ef_INTC_MSFT_RF$sd, ef_INTC_MSFT_RF$er, col="blue", pch=16, type = 'l', 
     xlab = "Standard Deviation", ylab = "Expected Return")
points(gmin.port.INTC_MSFT_RF$sd, gmin.port.INTC_MSFT_RF$er, col="red", 
       pch=16, cex=2)
lines(ef_INTC_MSFT_RF$sd[1:15], ef_INTC_MSFT_RF$er[1:15], col="green", pch=16, 
      type = 'l')
points(tan.port$sd, tan.port$er, col="magenta", pch=16, cex=2)
text(gmin.port.INTC_MSFT_RF$sd, gmin.port.INTC_MSFT_RF$er, 
     labels="GLOBAL MIN", pos=4)
text(tan.port$sd, tan.port$er, labels="TANGENCY", pos=4)
sr.tan = (tan.port$er - RF_mean)/tan.port$sd
abline(a=RF_mean, b=sr.tan, col="orange", lwd=2)


#All stocks with RF

ef_All_RF <- efficient.frontier(c(MSFT_mean,INTC_mean,JNJ_mean, LUV_mean, 
                                  MCD_mean), Cov_All)

# tangency portfolio
tan.port.all <- tangency.portfolio(c(MSFT_mean,INTC_mean,JNJ_mean, LUV_mean, 
                                     MCD_mean), Cov_All, RF_mean)

# compute global minimum variance portfolio
gmin.port.All_RF = globalMin.portfolio(c(MSFT_mean,INTC_mean,JNJ_mean, LUV_mean,
                                         MCD_mean), Cov_All)
plot(ef_All_RF$sd, ef_All_RF$er, col="blue", pch=16, type = 'l', 
     xlab = "Standard Deviation", ylab = "Expected Return", xlim = c(0,0.35),
     ylim = c(0, 0.27))
points(gmin.port.All_RF$sd, gmin.port.All_RF$er, col="red", pch=16, cex=2)
lines(ef_All_RF$sd[1:15], ef_All_RF$er[1:15], col="green", pch=16, type = 'l')
points(tan.port.all$sd, tan.port.all$er, col="magenta", pch=16, cex=2)
text(gmin.port.All_RF$sd, gmin.port.All_RF$er, labels="GLOBAL MIN", pos=1)
text(tan.port.all$sd, tan.port.all$er, labels="TANGENCY", pos=3)
sr.tan.all = (tan.port.all$er - RF_mean)/tan.port.all$sd
abline(a=RF_mean, b=sr.tan.all, col="orange", lwd=2)


gmin.port.All_RF = globalMin.portfolio(c(MSFT_mean,INTC_mean,JNJ_mean, LUV_mean,
                                         MCD_mean), Cov_All)
plot(ef_All_RF$sd, ef_All_RF$er, col="blue", pch=16, type = 'l', 
     xlab = "Standard Deviation", ylab = "Expected Return")
points(gmin.port.All_RF$sd, gmin.port.All_RF$er, col="red", pch=16, cex=2)
lines(ef_All_RF$sd[1:15], ef_All_RF$er[1:15], col="green", pch=16, type = 'l')
points(tan.port.all$sd, tan.port.all$er, col="magenta", pch=16, cex=2)
text(gmin.port.All_RF$sd, gmin.port.All_RF$er, labels="GLOBAL MIN", pos=4)
text(tan.port.all$sd, tan.port.all$er, labels="TANGENCY", pos=4)
sr.tan.all = (tan.port.all$er - RF_mean)/tan.port.all$sd
abline(a=RF_mean, b=sr.tan.all, col="orange", lwd=2)

```

Sharpe Ratio for MSFT-INTC portfolio: 0.773
Sharpe Ratio for portfolio with all stocks: 0.95

We can observe that the Sharpe Ratio improved as we kept adding more stocks (as we diversify). 



5. Assume your risk aversion is A = 4. What is your optimal mix of assets (including the
risk-free asset)?

```{r 5, echo=TRUE}

#Compute w*

w <- (tan.port.all$er - RF_mean)/(4*tan.port.all$sd^2)
```



6. Regress excess stock returns on excess market returns to obtain estimates of the betas
of the five stocks. Compute the standard errors of your estimates.

```{r 6, echo=TRUE}


#Microsoft
out_MSFT <- lm((MSFT_return-RF_return) ~ MktRF_return)
summary(out_MSFT)

#Intel
out_INTC <- lm((INTC_return-RF_return) ~ MktRF_return)
summary(out_INTC)

#Microsoft
out_JNJ <- lm((JNJ_return-RF_return) ~ MktRF_return)
summary(out_JNJ)

#Microsoft
out_LUV <- lm((LUV_return-RF_return) ~ MktRF_return)
summary(out_LUV)

#Microsoft
out_MCD <- lm((MCD_return-RF_return) ~ MktRF_return)
summary(out_MCD)

```
Estimate of Betas for MSFT, INTC, JNJ, LUV, MCD; in order: 0.9765890, 1.194556, 0.5633166, 1.080950, 0.6431096



7. What are the estimates of the alphas, and the standard deviation of these estimates?
What are the estimates of the standard deviation of idiosyncratic risk?

Estimate of the standard deviation of idiosyncratic risk is just same as the standard error of regression.

Estimate of Alpha for MSFT: 0.0027019
Standard deviation of Alpha for MSFT: 0.0008434
Estimates of the standard deviation of idiosyncratic risk for MSFT: 0.03368

Estimate of Alpha for INTC: 0.001451
Standard deviation of Alpha for INTC: 0.001008
Estimates of the standard deviation of idiosyncratic risk for INTC: 0.04025

Estimate of Alpha for JNJ: 0.0013050
Standard deviation of Alpha for JNJ: 0.0006293
Estimates of the standard deviation of idiosyncratic risk for JNJ: 0.02512

Estimate of Alpha for LUV: 0.001155
Standard deviation of Alpha for LUV: 0.001006
Estimates of the standard deviation of idiosyncratic risk for LUV: 0.04015

Estimate of Alpha for MCD: 0.0013048
Standard deviation of Alpha for MCD: 0.0006855
Estimates of the standard deviation of idiosyncratic risk for MCD: 0.02737



8. Compute the sample average excess return and compare the value with the return
predicted by the CAPM. Based on the data, how well does the CAPM predict the
level returns? How well does the CAPM predict relative performance?



```{r 8, echo=TRUE}

MSFT_mean - RF_mean
0.9765890*MktRF_mean

INTC_mean - RF_mean
1.194556*MktRF_mean

JNJ_mean - RF_mean
0.5633166*MktRF_mean

LUV_mean - RF_mean
1.080950*MktRF_mean

MCD_mean - RF_mean
0.6431096*MktRF_mean


```
Sample Average Excess Return of MSFT, INTC, JNJ, LUV, MCD; in order:  0.2241821, 0.1778056, 0.1161297, 0.1527046, 0.122957
Return Predicted by CAPM for MSFT, INTC, JNJ, LUV, MCD; in order: 0.08368289, 0.1023603, 0.04827001, 0.09262547, 0.05510739

Based on the data, we can see that CAPM does not properly predict the level returns and relative performance.  


