---
title: "Option Pricing"
description: "This project aims to create multiple option pricing models using different techniques for different use cases"
output:
  html_document:
    df_print: paged
  prettydoc::html_pretty:
    theme: architect
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(knitr)    
library(png)  
library(prettydoc)
```


## 1 Pricing European Options

```{r, results='asis'}
European_call <- function(S0, K, r, h, N, sigma, dividend) {
    dividend <- ifelse(dividend<0, 0, dividend)
    u <- exp(((r-dividend)*h) + (sigma*sqrt(h)))
    d <- exp(((r-dividend)*h) - (sigma*sqrt(h)))
    S <- vector(mode = "list", length = N+1);
    S[[N+1]][1] <- 0
    option <- vector(mode = "list", length = N+1);
    option[[N+1]][1] <- 0;
    delta <- vector(mode = "list", length = N+1);
    delta[[N+1]][1] <- 0;
    B <- vector(mode = "list", length = N+1);
    B[[N+1]][1] <- 0;
    S[[1]][1] <- S0;
    for (i in (N+1):1) {
        for (j in 1:i) {
            S[[i]][j] <- S0 * (u^(i-j))*(d^(j-1));
            if(i == N+1) { 
                option[[i]][j] <- max(S[[i]][j]-K, 0);
                delta[[i]][j] <- 0;
                B[[i]][j] <- 0;
            } else {
                option[[i]][j] <- exp(-r*h) * ((option[[i+1]][j] * (exp((r-dividend)*h)-d) / (u-d)) + 
                                                   (option[[i+1]][j+1] * (u-exp((r-dividend)*h)) / (u-d)));
                delta[[i]][j] <- (option[[i+1]][j] - option[[i+1]][j+1]) / (S[[i]][j]*(u-d));
                B[[i]][j] <- exp(-r*h) * (((option[[i+1]][j+1]*u) - (option[[i+1]][j]*d)) / (u-d));
                
            }
        }
    }
    p <- (exp(r*h)-d) / (u-d)
    S_v <- S[[1]]
    option_v <- option[[1]]
    delta_v <- delta[[1]]
    B_v <- B[[1]]
    T_v <- 0
    for (i in 2:(N+1)) {
        S_v <- S_v %>% append(S[[i]])
        option_v <- option_v %>% append(option[[i]])
        delta_v <- delta_v %>% append(delta[[i]])
        B_v <- B_v %>% append(B[[i]])
        T_v <- T_v %>% append(rep(h*(i-1), times=i))
    }
    output <- data.frame(T_v, S_v, option_v, delta_v, B_v, "u" = rep(u, times = length(B_v)), 
                         "d" = rep(d, times = length(B_v)), "p_star" = rep(p, times = length(B_v)))
    return(output)
}

European_put <- function(S0, K, r, h, N, sigma, dividend) {
    dividend <- ifelse(dividend<0, 0, dividend)
    u <- exp(((r-dividend)*h) + (sigma*sqrt(h)))
    d <- exp(((r-dividend)*h) - (sigma*sqrt(h)))
    S <- vector(mode = "list", length = N+1);
    S[[N+1]][1] <- 0
    option <- vector(mode = "list", length = N+1);
    option[[N+1]][1] <- 0;
    delta <- vector(mode = "list", length = N+1);
    delta[[N+1]][1] <- 0;
    B <- vector(mode = "list", length = N+1);
    B[[N+1]][1] <- 0;
    S[[1]][1] <- S0;
    for (i in (N+1):1) {
        for (j in 1:i) {
            S[[i]][j] <- S0 * (u^(i-j))*(d^(j-1));
            if(i == N+1) { 
                option[[i]][j] <- max(K-S[[i]][j], 0);
                delta[[i]][j] <- 0;
                B[[i]][j] <- 0;
            } else {
                option[[i]][j] <- exp(-r*h) * ((option[[i+1]][j] * (exp((r-dividend)*h)-d) / (u-d)) + 
                                                   (option[[i+1]][j+1] * (u-exp((r-dividend)*h)) / (u-d)));
                delta[[i]][j] <- (option[[i+1]][j] - option[[i+1]][j+1]) / (S[[i]][j]*(u-d));
                B[[i]][j] <- exp(-r*h) * (((option[[i+1]][j+1]*u) - (option[[i+1]][j]*d)) / (u-d));
                
            }
        }
    }
    p <- (exp(r*h)-d) / (u-d)
    S_v <- S[[1]]
    option_v <- option[[1]]
    delta_v <- delta[[1]]
    B_v <- B[[1]]
    T_v <- 0
    for (i in 2:(N+1)) {
        S_v <- S_v %>% append(S[[i]])
        option_v <- option_v %>% append(option[[i]])
        delta_v <- delta_v %>% append(delta[[i]])
        B_v <- B_v %>% append(B[[i]])
        T_v <- T_v %>% append(rep(h*(i-1), times=i))
    }
    output <- data.frame(T_v, S_v, option_v, delta_v, B_v, "u" = rep(u, times = length(B_v)), 
                         "d" = rep(d, times = length(B_v)), "p_star" = rep(p, times = length(B_v)))
    return(output)
}

e_straddle <- function(S0, K, r, h, N, sigma, dividend) {
    dividend <- ifelse(dividend<0, 0, dividend)
    u <- exp(((r-dividend)*h) + (sigma*sqrt(h)))
    d <- exp(((r-dividend)*h) - (sigma*sqrt(h)))
    S <- vector(mode = "list", length = N+1);
    S[[N+1]][1] <- 0
    option <- vector(mode = "list", length = N+1);
    option[[N+1]][1] <- 0;
    delta <- vector(mode = "list", length = N+1);
    delta[[N+1]][1] <- 0;
    B <- vector(mode = "list", length = N+1);
    B[[N+1]][1] <- 0;
    S[[1]][1] <- S0;
    for (i in (N+1):1) {
        for (j in 1:i) {
            S[[i]][j] <- S0 * (u^(i-j))*(d^(j-1));
            if(i == N+1) { 
                option[[i]][j] <- (max(S[[i]][j]-K, 0) + max(K-S[[i]][j], 0));
                delta[[i]][j] <- 0;
                B[[i]][j] <- 0;
            } else {
                option[[i]][j] <- exp(-r*h) * ((option[[i+1]][j] * (exp((r-dividend)*h)-d) / (u-d)) + 
                                                       (option[[i+1]][j+1] * (u-exp((r-dividend)*h)) / (u-d)));
                delta[[i]][j] <- (option[[i+1]][j] - option[[i+1]][j+1]) / (S[[i]][j]*(u-d));
                B[[i]][j] <- exp(-r*h) * (((option[[i+1]][j+1]*u) - (option[[i+1]][j]*d)) / (u-d));
            }
        }
    }
    p <- (exp(r*h)-d) / (u-d)
    S_v <- S[[1]]
    option_v <- option[[1]]
    delta_v <- delta[[1]]
    B_v <- B[[1]]
    T_v <- 0
    for (i in 2:(N+1)) {
        S_v <- S_v %>% append(S[[i]])
        option_v <- option_v %>% append(option[[i]])
        delta_v <- delta_v %>% append(delta[[i]])
        B_v <- B_v %>% append(B[[i]])
        T_v <- T_v %>% append(rep(h*(i-1), times=i))
    }
    output <- data.frame(T_v, S_v, option_v, delta_v, B_v, "u" = rep(u, times = length(B_v)), 
                         "d" = rep(d, times = length(B_v)), "p_star" = rep(p, times = length(B_v)))
    return(output)
}
```

1. 
```{r}
e_straddle(100, 90, 0.02, 0.25, 4, 0.2, 0)
```

2. 
```{r}
e_straddle(100, 90, 0.02, 0.025, 40, 0.2, 0)
```

3.
```{r}
European_call(100, 90, 0.02, 0.25, 4, 0.2, 0)
```

## 2 Pricing American Options

```{r, results='asis'}
American_call <- function(S0, K, r, h, N, sigma, dividend) {
    dividend <- ifelse(dividend<0, 0, dividend)
    u <- exp(((r-dividend)*h) + (sigma*sqrt(h)))
    d <- exp(((r-dividend)*h) - (sigma*sqrt(h)))
    S <- vector(mode = "list", length = N+1);
    S[[N+1]][1] <- 0
    ex <- vector(mode = "list", length = N+1);
    ex[[N+1]][1] <- "Not Executed Early"
    option <- vector(mode = "list", length = N+1);
    option[[N+1]][1] <- 0;
    delta <- vector(mode = "list", length = N+1);
    delta[[N+1]][1] <- 0;
    B <- vector(mode = "list", length = N+1);
    B[[N+1]][1] <- 0;
    S[[1]][1] <- S0;
    for (i in (N+1):1) {
        for (j in 1:i) {
            S[[i]][j] <- S0 * (u^(i-j))*(d^(j-1));
            if(i == N+1) { 
                option[[i]][j] <- max(S[[i]][j]-K, 0);
                delta[[i]][j] <- 0;
                B[[i]][j] <- 0;
                ex[[i]][j] <- "Expiration";
            } else {
                if(exp(-r*h) * ((option[[i+1]][j] * (exp((r-dividend)*h)-d) / (u-d)) + 
                                (option[[i+1]][j+1] * (u-exp((r-dividend)*h)) / (u-d))) >=
                   max(S[[i]][j]-K, 0)) {
                    option[[i]][j] <- exp(-r*h) * ((option[[i+1]][j] * (exp((r-dividend)*h)-d) / (u-d)) + 
                                                       (option[[i+1]][j+1] * (u-exp((r-dividend)*h)) / (u-d)));
                    ex[[i]][j] <- "Not Executed Early";
                } else {
                    option[[i]][j] <- max(S[[i]][j]-K, 0);
                    ex[[i]][j] <- "Executed Early";
                }
                delta[[i]][j] <- (option[[i+1]][j] - option[[i+1]][j+1]) / (S[[i]][j]*(u-d));
                B[[i]][j] <- exp(-r*h) * (((option[[i+1]][j+1]*u) - (option[[i+1]][j]*d)) / (u-d));
                
            }
        }
    }
    p <- (exp(r*h)-d) / (u-d)
    ex_v <- ex[[1]]
    S_v <- S[[1]]
    option_v <- option[[1]]
    delta_v <- delta[[1]]
    B_v <- B[[1]]
    T_v <- 0
    for (i in 2:(N+1)) {
        ex_v <- ex_v %>% append(ex[[i]])
        S_v <- S_v %>% append(S[[i]])
        option_v <- option_v %>% append(option[[i]])
        delta_v <- delta_v %>% append(delta[[i]])
        B_v <- B_v %>% append(B[[i]])
        T_v <- T_v %>% append(rep(h*(i-1), times=i))
    }
    output <- data.frame(T_v, S_v, option_v, delta_v, B_v, ex_v, "u" = rep(u, times = length(B_v)), 
                         "d" = rep(d, times = length(B_v)), "p_star" = rep(p, times = length(B_v)))
    return(output)
}

American_call(10, 10, 0.01, 1/365, 250, 0.15, 0)
```

## 3 Discrete Dividends

```{r}
American_call <- function(S0, K, r, h, N, sigma, dividend, dt) {
    dividend <- ifelse(dividend<0, 0, dividend)
    u <- exp(sigma*sqrt(h))
    d <- 1/u
    S <- vector(mode = "list", length = N+1);
    S[[N+1]][1] <- 0
    ex <- vector(mode = "list", length = N+1);
    ex[[N+1]][1] <- "Not Executed Early"
    option <- vector(mode = "list", length = N+1);
    option[[N+1]][1] <- 0;
    delta <- vector(mode = "list", length = N+1);
    delta[[N+1]][1] <- 0;
    B <- vector(mode = "list", length = N+1);
    B[[N+1]][1] <- 0;
    S[[1]][1] <- S0;
    for (i in 2:(N+1)) {
        for (j in 1:i) {
            if (i == j) {
                if (i %in% dt) {
                    S[[i]][j] <- (1 - dividend) * S[[i-1]][j-1] * d;
                } else {
                     S[[i]][j] <- S[[i-1]][j-1] * d;
                }
            } else {
                if (i %in% dt) {
                    S[[i]][j] <- (1 - dividend) * S[[i-1]][j] * u;
                } else {
                     S[[i]][j] <- S[[i-1]][j] * u;
                }
            }
        }
    }
    for (i in (N+1):1) {
        for (j in 1:i) {
            if(i == N+1) { 
                option[[i]][j] <- max(S[[i]][j]-K, 0);
                delta[[i]][j] <- 0;
                B[[i]][j] <- 0;
                ex[[i]][j] <- "Expiration";
            } else {
                if(exp(-r*h) * ((option[[i+1]][j] * (exp((r-dividend)*h)-d) / (u-d)) +
                                (option[[i+1]][j+1] * (u-exp((r-dividend)*h)) / (u-d))) >=
                   max(S[[i]][j]-K, 0)) {
                    option[[i]][j] <- exp(-r*h) * ((option[[i+1]][j] * (exp((r-dividend)*h)-d) / (u-d)) +
                                                       (option[[i+1]][j+1] * (u-exp((r-dividend)*h)) / (u-d)));
                    ex[[i]][j] <- "Not Executed Early";
                } else {
                    option[[i]][j] <- max(S[[i]][j]-K, 0);
                    ex[[i]][j] <- "Executed Early";
                }
                delta[[i]][j] <- (option[[i+1]][j] - option[[i+1]][j+1]) / (S[[i]][j]*(u-d));
                B[[i]][j] <- exp(-r*h) * (((option[[i+1]][j+1]*u) - (option[[i+1]][j]*d)) / (u-d));
            }
        }
    }
    return(option[[1]])
}

American_put <- function(S0, K, r, h, N, sigma, dividend, dt) {
    dividend <- ifelse(dividend<0, 0, dividend)
    u <- exp(((r-dividend)*h) + (sigma*sqrt(h)))
    d <- exp(((r-dividend)*h) - (sigma*sqrt(h)))
    S <- vector(mode = "list", length = N+1);
    S[[N+1]][1] <- 0
    ex <- vector(mode = "list", length = N+1);
    ex[[N+1]][1] <- "Not Executed Early"
    option <- vector(mode = "list", length = N+1);
    option[[N+1]][1] <- 0;
    delta <- vector(mode = "list", length = N+1);
    delta[[N+1]][1] <- 0;
    B <- vector(mode = "list", length = N+1);
    B[[N+1]][1] <- 0;
    S[[1]][1] <- S0;
    for (i in 2:(N+1)) {
        for (j in 1:i) {
            if (i == j) {
                if (i %in% dt) {
                    S[[i]][j] <- (1 - dividend) * S[[i-1]][j-1] * d;
                } else {
                     S[[i]][j] <- S[[i-1]][j-1] * d;
                }
            } else {
                if (i %in% dt) {
                    S[[i]][j] <- (1 - dividend) * S[[i-1]][j] * u;
                } else {
                     S[[i]][j] <- S[[i-1]][j] * u;
                }
            }
        }
    }
    for (i in (N+1):1) {
        for (j in 1:i) {
            if(i == N+1) { 
                option[[i]][j] <- max(K-S[[i]][j], 0);
                delta[[i]][j] <- 0;
                B[[i]][j] <- 0;
                ex[[i]][j] <- "Expiration";
            } else {
                if(exp(-r*h) * ((option[[i+1]][j] * (exp((r-dividend)*h)-d) / (u-d)) + 
                                (option[[i+1]][j+1] * (u-exp((r-dividend)*h)) / (u-d))) >=
                   max(K-S[[i]][j], 0)) {
                    option[[i]][j] <- exp(-r*h) * ((option[[i+1]][j] * (exp((r-dividend)*h)-d) / (u-d)) + 
                                                       (option[[i+1]][j+1] * (u-exp((r-dividend)*h)) / (u-d)));
                    ex[[i]][j] <- "Not Executed Early";
                } else {
                    option[[i]][j] <- max(K-S[[i]][j], 0);
                    ex[[i]][j] <- "Executed Early";
                }
                delta[[i]][j] <- (option[[i+1]][j] - option[[i+1]][j+1]) / (S[[i]][j]*(u-d));
                B[[i]][j] <- exp(-r*h) * (((option[[i+1]][j+1]*u) - (option[[i+1]][j]*d)) / (u-d));
                
            }
        }
    }
    return(option[[1]])
}


straddle <- function(S0, K, r, h, N, sigma, dividend, dt) {
    dividend <- ifelse(dividend<0, 0, dividend)
    u <- exp(((r-dividend)*h) + (sigma*sqrt(h)))
    d <- exp(((r-dividend)*h) - (sigma*sqrt(h)))
    S <- vector(mode = "list", length = N+1);
    S[[N+1]][1] <- 0
    option <- vector(mode = "list", length = N+1);
    option[[N+1]][1] <- 0;
    delta <- vector(mode = "list", length = N+1);
    delta[[N+1]][1] <- 0;
    B <- vector(mode = "list", length = N+1);
    B[[N+1]][1] <- 0;
    S[[1]][1] <- S0;
    for (i in 2:(N+1)) {
        for (j in 1:i) {
            if (i == j) {
                if (i %in% dt) {
                    S[[i]][j] <- (1 - dividend) * S[[i-1]][j-1] * d;
                } else {
                     S[[i]][j] <- S[[i-1]][j-1] * d;
                }
            } else {
                if (i %in% dt) {
                    S[[i]][j] <- (1 - dividend) * S[[i-1]][j] * u;
                } else {
                     S[[i]][j] <- S[[i-1]][j] * u;
                }
            }
        }
    }
    for (i in (N+1):1) {
        for (j in 1:i) {
            if(i == N+1) { 
                option[[i]][j] <- (max(S[[i]][j]-K, 0) + max(K-S[[i]][j], 0));
                delta[[i]][j] <- 0;
                B[[i]][j] <- 0;
            } else {
                if(exp(-r*h) * ((option[[i+1]][j] * (exp((r-dividend)*h)-d) / (u-d)) + 
                                (option[[i+1]][j+1] * (u-exp((r-dividend)*h)) / (u-d))) >=
                   (max(S[[i]][j]-K, 0) + max(K-S[[i]][j], 0))) {
                    option[[i]][j] <- exp(-r*h) * ((option[[i+1]][j] * (exp((r-dividend)*h)-d) / (u-d)) + 
                                                       (option[[i+1]][j+1] * (u-exp((r-dividend)*h)) / (u-d)));
                } else {
                    option[[i]][j] <- (max(S[[i]][j]-K, 0) + max(K-S[[i]][j], 0));
                }
                delta[[i]][j] <- (option[[i+1]][j] - option[[i+1]][j+1]) / (S[[i]][j]*(u-d));
                B[[i]][j] <- exp(-r*h) * (((option[[i+1]][j+1]*u) - (option[[i+1]][j]*d)) / (u-d));
            }
        }
    }
    return(option[[1]])
}
```

1. 
```{r}
American_call(10, 10, 0.02, 1/365, 200, 0.2, 0.05, c(50, 100, 150))

American_put(10, 10, 0.02, 1/365, 200, 0.2, 0.05, c(50, 100, 150))
```

2. 
```{r}
straddle(10, 10, 0.02, 1/365, 200, 0.2, 0.05, c(50, 100, 150))
```
Since the American call should never be exercised, the optimal exercise rule of the American straddle implies that upon exercising you lose the difference between the value of the call and its intrinsic value. These loses percolate back to the initial node and imply that the American straddle is worth less than the portfolio of American options.

## 4 Pricing Asian Options by Monte Carlo

```{r}
Asian_Option <- function(s,k,r,t,n,div=0,sd=0){
  
  S <- 0
  C <- 0
  h <- t/n
  
for(j in 1:1000){
  S <- 0
  for (i in 1:n) {
    if(i==1){
      S[i] <- s*exp((r-((sd)^2/2))*h + (sd*sqrt(h)*rnorm(1,0,1)))
    }else{
      S[i] <- S[i-1]*exp((r-((sd)^2/2))*h + (sd*sqrt(h)*rnorm(1,0,1)))
    }
  }
  C[j] <- max(mean(S)-k,0)
}
  return(mean(C)*exp(-r))
}

sd <- 0.2
Asian_Call <- Asian_Option(200,220,0.02,1,365,0,0.2)
Asian_Call
Asian_Call_ConfInt_Up <- Asian_Call + qnorm(0.975)*(sd/sqrt(100000))
Asian_Call_ConfInt_Up
Asian_Call_ConfInt_Down <- Asian_Call - qnorm(0.975)*(sd/sqrt(100000))
Asian_Call_ConfInt_Down
```

## 5 Pricing American Options using the Longstaff and Schwartz LeastSquare Method

```{r}

S0 <- 200
K <- 220

generate_monte_carlo<-function(N, paths)
{
  interval<-1/N
  mean <- 0.06
  sigma <- 0.2
  S<-matrix(0,nrow=(N+1),ncol=paths)
  for(i in 1:paths)
  {
    S[1,i]<- S0
    for(j in 2:(N+1))
    {
      S[j,i]<-S[j-1,i]*exp(interval*(mean-((sigma)^2)/2)+((interval)^.5)*rnorm(1)*sigma)
    }
  }
  return(S)
  }

longstaff_schwartz <- function(ey, x, N){
  
  ey <- exp(-0.1/N) * ey
  df <- data.frame(x, x2 = x^2, ey)
  out <- lm(ey ~ x + x2, data = df)
  return(out)
}

calculate_call_price <- function(N, paths){
C<-matrix(0,nrow=N,ncol=paths)
S <- generate_monte_carlo(N, paths) 
calllist <- list()
pricelist <- list()
prevcallist <- list()


for (i in N:2) {
  for (j in 1:paths) {
    if (i == N) {
    C[i,j] = max(K - S[i+1,j], 0)
    }
    C[i-1,j] = max(K - S[i,j], 0)
  }
  for (j in 1:paths) {
    if (C[i-1, j] > 0) {
    calllist[j] <- C[i-1, j]
    pricelist[j] <- S[i, j]
    prevcallist[j] <- C[i, j]
    }
  }
  calllist <- calllist[!sapply(calllist,is.null)]
  pricelist <- pricelist[!sapply(pricelist,is.null)]
  prevcallist <- prevcallist[!sapply(prevcallist,is.null)]

  calldf <- t(as.data.frame(calllist))
  pricedf <- t(as.data.frame(pricelist))
  prevcalldf <- t(as.data.frame(prevcallist))
  outdf <- longstaff_schwartz(prevcalldf, pricedf, N)
  
  for (j in 1:paths) {
    if (C[i-1, j] > (outdf$coefficients[1] + outdf$coefficients[2] * S[i,j] + outdf$coefficients[3] * S[i,j]^2)){
      C[i:N,j] <- 0
    }else{
      C[i-1, j] <- 0
    }
  }
  
}
total <- list()
for (i in 1:N) {
  total[i] <- sum(C[i,]) * exp(-i*0.1/N)
}

total <- total[!sapply(total,is.null)]
total <- t(as.data.frame(total))
call_price <- sum(total)/paths
call_price
}
```

1. N = 250 and paths = 100000.

```{r}
#calculate_call_price(250, 100000)
```

2. N = 250 and paths = 10, 100, 1000, 10000, 100000.  

```{r}
calculate_call_price(250, 10)
calculate_call_price(250, 100)
calculate_call_price(250, 1000)
#calculate_call_price(250, 10000)
#calculate_call_price(250, 100000)
```

3. N = 3, 10, 100, 250, 100000 and paths = 100000.

```{r}
calculate_call_price(3, 100000)
#calculate_call_price(10, 100000)
#calculate_call_price(100, 100000)
#calculate_call_price(250, 100000)
#calculate_call_price(1000, 100000)
```
PS: Please note that in the above, the actual runs are not performed for paths = 10000 and 100000 in sections 1 & 2 and for N = 10, 100, 250, 1000 in section 3 due to long run times. 


